### 2014-05-10 text/Chapter04.tex
Start working on Chapter 4

\chapter{Flat coeffect language} 
\label{ch:flat}

>>>>

%\begin{flushright}{\slshape    
%Even the most puritanical rationalist will be forced to stop arguing and \\
%use propaganda [...] because the psychological conditions have disappeared \\
%that allowed effective argument and therefore influence over the others.} \\ \medskip
%--- Paul Feyerabend, \emph{Against Method} \cite{philosophy-feyerabend}
%\end{flushright}
%\vspace{2em}

\chapter{Flat coeffect calculus} 
\label{ch:flat}

Successful programming language abstractions need to generalize a wide range of recurring
problems while capturing the key commonalities. These two aims are typically in opposition -- 
more general abstractions provide smaller value, while less general abstractions cannot be
used as often.

In the previous chapter, we outlined a number of systems that capture how computations
access the environment in which they are executed. We identified two kinds of systems --
\emph{flat} capturing whole-context properties and \emph{structural} capturing per-variable
properties. As we show in Chapter~\ref{ch:unified}, the systems can be unified using a single 
abstraction. This is useful when implementing and composing the systems, but such abstraction 
provides \emph{smaller value} -- \ie~its generality hides useful properties that we can see 
when we consider the systems separately. For this reason, this and the next chapter discusses 
\emph{flat} and \emph{structural} systems separately.

\section{Introduction}
\label{sec:flat-intro}

Liveness is more similar to data-flow.

Implicit parameters (or resources) are a bit different, but important.

The closest work is Uustalu and Vene

\section{Flat coeffect calculus}
\label{sec:flat-calculus}

This section revisits the flat coeffect calculus, which provides a 
basis for the structural coeffect calculus that we introduce in Section~... We 
make a number of improvements -- we cover more standard language features and require additional laws
which allows a more useful equational theory.
The formulation is general, so as not to restrict the possible uses unnecessarily, but is
specialised enough to provide useful properties.

The flat coeffect calculus maintains a single annotation of the contextual requirements associated
with the variable context. It can be instantiated to capture implicit parameters, the system for 
dataflow and simple liveness analysis.

The expressions are those of the $\lambda$-calculus with \emph{let} binding; assuming $T$ ranges
over base types, types are extended as follows:
%
\begin{equation*}
\begin{array}{rcl}
e &::=& x \sep \lambda x.e \sep e_1~e_2 \sep \kvd{let}~x = e_1~\kvd{in}~e_2\\
\tau &::=& T \sep \tau_1 \xrightarrow{\cclrd{r}} \tau_2
\end{array}
\end{equation*}
%
where $T$ ranges over type constants. 
We discuss pairs and recursion in Sections~\ref{sec:flat-synprop} and \ref{sec:structural-properties}.
The type $\tau_1 \xrightarrow{\cclrd{r}} \tau_2$
represents a function from $\tau_1$ to $\tau_2$ that requires additional context $\cclrd{r}$.
It can be viewed as a pure function that takes $\tau_1$ \emph{with} context $r$.

%Note that
%$\ctxtyp{r}{\tau}$ does not exist as syntactic value (although it has a semantic meaning
%discussed in Section~\ref{sec:flat-semantics}). The annotations $r$ are formed by 
%an algebraic structure discussed next.

\subsection{Flat coeffect algebra}

\begin{definition}
A \emph{\cclrd{flat coeffect algebra}} $(\C, \cseq, \cpar, \czip, \cunit, \czero, \cleq)$ is a set 
$\C$ together with elements $\cunit, \czero \in \C$, relation $\cleq$ and binary operations 
$\cseq, \cpar, \czip$ such that $(\C, \cseq, \cunit)$ and $(\C, \cpar, \czero)$ are monoids,
$(\C, \cleq)$ is a pre-order and $(\C, \czip)$ is a semigroup. That is, for all $r,s,t\in \C$:
%
\begin{equation*}
\begin{array}{ccr}
r \;\cseq\; (s \;\cseq\; t) = (r \;\cseq\; s) \;\cseq\; t  &
\cunit \;\cseq\; r = r = r \;\cseq\; \cunit &
\textnormal{(monoid)}   
\\
r \;\cpar\; (s \;\cpar\; t) = (r \;\cpar\; s) \;\cpar\; t &
\czero \;\cpar\; r = r = r \;\cpar\; \czero &
\textnormal{(monoid)}   
\\
\textnormal{if}~~r\; \cleq\; s ~~\textnormal{and}~~s\; \cleq\; t~~\textnormal{then}~~r\; \cleq\; t&
t\; \cleq\; t &
\textnormal{(pre-order)}   
\\
r \;\czip\; (s \;\czip\; t) = (r \;\czip\; s) \;\czip\; t &&
\textnormal{(assoc)}   
\end{array}
\end{equation*}
%
In addition, the following distributivity axioms hold:
\begin{align*}
\quad (\cclrd{r}\, \cpar\, \cclrd{s}) \;\cseq\; \cclrd{t} & = (\cclrd{r} \,\cseq\, \cclrd{t}) \;\cpar\; (\cclrd{s}\, \cseq\, \cclrd{t}) \\
\quad \cclrd{t} \;\cseq\; (\cclrd{r}\, \cpar\, \cclrd{s}) & = (\cclrd{t} \,\cseq\, \cclrd{r}) \;\cpar\; (\cclrd{t}\, \cseq\, \cclrd{s})
\end{align*}
\end{definition}

\subsection{Flat coeffect types}

\subsection{Examples of flat coeffects}

\section{Categorical motivation}
\label{sec:flat-semantics}

\subsection{Comonads}
\subsection{Indexed comonads}
\subsection{Semantics of flat calculus}
\subsection{Examples}

\section{Equational theory}
\label{sec:flat-syntax}

\subsection{Call-by-value evaluation}
\subsection{Call-by-name evaluation}

\section{Syntactic extensions}
\label{sec:flat-exts}

\subsection{Lambda abstraction}
\subsection{Constants and pairs}
\subsection{Recursion}

\section{Type inference}
\label{sec:flat-inference}

\subsection{Semi-lattice formulation}
\subsection{Type inference algorithm}

\section{Summary}

\section{----------------JUNK!}

\begin{figure}[t]
\begin{equation*}
\tyrule{var}
  {x : \tau \in \Gamma}
  {\coctx{\Gamma}{\aclrd{ 0 }} \vdash x : \tau }
\end{equation*}
\begin{equation*}
\tyrule{sub}
  {\coctx{\Gamma}{\aclrd{n'}} \vdash e : \tau }
  {\coctx{\Gamma}{\aclrd{n}} \vdash e : \tau }\quad\quad(\aclrd{n'} \leq \aclrd{n})
\end{equation*}
\begin{equation*}
\tyrule{app}
  {\coctx{\Gamma}{\aclrd{m}} \vdash e_1 : \tau_1 \xrightarrow{\aclrd{p}} \tau_2 &
   \coctx{\Gamma}{\aclrd{n}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma}{\textnormal{max}(\aclrd{m}, \aclrd{n} + \aclrd{p})} \vdash e_1~e_2 : \tau_2}
\end{equation*}
\begin{equation*}
\tyrule{let}
  { \coctx{\Gamma}{\aclrd{m}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma, x:\tau_1}{\aclrd{n}} \vdash e_2 : \tau_2}
  {\coctx{\Gamma}{\aclrd{n}+\aclrd{m}} \vdash \kvd{let}~x=e_1~\kvd{in}~e_2 : \tau_2 }
\end{equation*}
\begin{equation*}
\tyrule{abs}
  {\coctx{\Gamma, x:\tau_1}{\aclrd{n}} \vdash e : \tau_2}
  {\coctx{\Gamma}{\aclrd{n}} \vdash \lambda x.e : \tau_1 \xrightarrow{\aclrd{n}} \tau_2 }
\end{equation*}

\caption{Flat coeffect calculus}
\label{fig:flat-types}
\end{figure}

% ------------------

\section{Type inference}

~
\newpage
~
